| Environmental and Development Economics\n| Week 2 - How does development affect the environment?

### 2024-08-30 assignments/research_proposal/research_proposal.Rmd
update first week

@@

## Timeline
\begin{table}[ht]
\begin{tabular}{|c|c|}
\hline
\textbf{Item} & \textbf{Deadline} \\ \hline
First draft     & October 3rd     \\ \hline
Referee report    & October 10th     \\ \hline
Final draft    & October 31st     \\ \hline
\end{tabular}
\end{table}

@@

- Marc Bellemare's guide on \href{https://marcfbellemare.com/wordpress/wp-content/uploads/2020/09/BellemareHowToPaperSeptember2020.pdf}{how to write applied economics papers}
- Keith Head's \href{https://blogs.ubc.ca/khead/research/research-advice/formula}{introduction formula}
- Plamen Nikolov's \href{https://docs.iza.org/dp15057.pdf}{writing tips for economics research papers}
- Read papers!

>>>>

**Due date: Final draft due Oct. 31st, 2024**

@@

## Timeline
\begin{table}[ht]
\begin{tabular}{|c|c|}
\hline
\textbf{Item} & \textbf{Deadline} \\ \hline
First draft     & October 3rd     \\ \hline
Referee report    & October 10th     \\ \hline
Final draft    & October 31st     \\ \hline
\end{tabular}
\end{table}
Note: the first draft is pass/fail. I will check it for completion and award a pass. If you fail to submit a first draft, then five percentage points will be taken off of your final grade.

@@

- Marc Bellemare's guide on \href{https://marcfbellemare.com/wordpress/wp-content/uploads/2020/09/BellemareHowToPaperSeptember2020.pdf}{how to write applied economics papers}
- Keith Head's \href{https://blogs.ubc.ca/khead/research/research-advice/formula}{introduction formula}
- Plamen Nikolov's \href{https://docs.iza.org/dp15057.pdf}{writing tips for economics research papers}

### 2024-08-30 lectures/module1/module1.Rmd
update first week

---
title:  |
    | Environmental and Development Economics
    | Week 1 - Introduction
author:
  |
    | Raahil Madhok
    | UMN Applied Economics
date: "`r Sys.Date()`"

@@

# This includes latex arguments
header-includes:
  - \AtBeginDocument{\title[Week 1 - Introduction]{\textbf{Environmental and Development Economics} \\ Week 1 - Introduction}}
  - \input{header.tex}

@@

@@

- Why study environmental economics in LMICs?
\vfill
- Course overview + detailed outline
\vfill
- Grade breakdown
\vfill
- Course website: https://github.com/rmadhok/enviro-dev-grad
  - lectures, assignments, syllabus

@@

Instead, I am organizing around **\textcolor{orange}{FOUR}** key questions:

1. Why is environmental quality so bad in developing countries?

2. What are the costs of poor environmental quality in developing countries?

3. Why is WTP for environmental quality low in developing countries?

4. What are the political economy barriers to environmental protection?

@@

- Teach applied papers
  - research design, identification strategy, estimation techniques

- Emphasize recent JMPs

@@

## Weeks 1 and 2

**Week 1: Theory of Environment/Development**

- How to use theory to ask the right questions (lecture 2)

@@

**Week 2: The effect of development on the environment**

- air quality, water quality (lecture 3)
- forests, biodiversity (lecture 4)

## Weeks 3 and 4

**Week 3: The effect of the environment on development**

- human capital (lecture 5)
- firms (lecture 6)

**Week 4: WTP for Environmental Quality**

- TBD (lecture 7)
- TBD (lecture 8)

## Weeks 5, 6, and 7

**Week 5: Environmental policy design**

- Spillovers (lecture 9)
- Pollution markets in developing countries (lecture 10)

**Week 6: Political economy of the environment**

- Corruption and weak enforcement  (lecture 11)
- Environmental justice (lecture 12)

**Week 7: Research presentations**

- I will provide more details throughout the semester

@@

\begin{table}[]
\begin{tabular}{lllll}
In-class presentations      & 10\% &  &  &  \\
Problem Set           & 20\% &  &  &  \\
Research Proposal & 60 \% &  &  &  \\
Participation                     & 10\%     &  &  & 
\end{tabular}
\end{table}

@@

- At start of **first** lecture each week, you'll give a 15 min paper presentation

- Each student submits **seven** summary slides (5% of grade)
  - motivation, research question, methods, results
  - 10 mins presentation + 5 mins Q&A (5% of grade)

@@

- You will replicate an environment/development paper
  - You will also **extend** the results

@@

\begin{table}[]
\begin{tabular}{lllll}
Written Proposal      & 30\% &  &  &  \\
Peer Review           & 20\% &  &  &  \\
Proposal Presentation & 10\% &  &  &  \\
                      &      &  &  & 
\end{tabular}
\end{table}

- You will develop a research proposal for an original idea
  - You are NOT expected to actually do the analysis  
  - I will provide small deadlines (outline, first draft, etc.) along the way

@@

## Next week

@@

- Impact of development on the environment (air, water)

@@

>>>>

---
title:  |
    | Environmental and Development Economics
    | Module 1 - Introduction
author:
  |
    | Raahil Madhok
    | UMN Applied Economics
date: "`r Sys.Date()`"

@@

# This includes latex arguments
header-includes:
  - \AtBeginDocument{\title[Module 1 - Introduction]{\textbf{Environmental and Development Economics} \\ Module 1 - Introduction}}
  - \input{header.tex}

@@

## Housekeeping

- **Class Time/Location:** Tues/Thurs 11:45am-1:25pm, Ruttan 119
\vfill
- **Office Hours:** Thursdays 1:30pm-2:30pm, Ruttan 337D
\vfill
- **Course website:** https://github.com/rmadhok/enviro-dev-grad
  - lectures, assignments, syllabus
  - Try to skim reading(s) beforehand
\vfill
- **Assignments**: Upload through Canvas
\vfill

@@

- Why study environmental economics in LMICs?
\vfill
- Course overview + detailed outline
\vfill
- Grade breakdown
\vfill
- If time: conceptual framework for environment & development econ

@@

Instead, I am organizing around **\textcolor{orange}{FIVE}** key questions:

1. How does economic development affect the environment, and vice versa?

2. Why is environmental quality so bad in developing countries?

3. What are the costs of poor environmental quality in developing countries?

4. Why is WTP for environmental quality low in developing countries?

5. What are the political economy barriers to environmental protection?

@@

- Teach applied papers
  - research design, identification strategy, estimation techniques
  - aim for two per class (please skim beforehand)

- Emphasize recent papers

@@

## Course Outline

**Module 1: Introduction**

- Lecture 1: Course intro + how to use theory to ask the right questions

@@

**Module 2: The effect of development on the environment**

- Lecture 2: Income effects
- Lecture 3: Access to capital (technology and infrastructure)

## Course Outline

**Module 3: The effect of environment on development**

- Lecture 4: Health
- Lecture 5: Productivity
\vfill

**Module 4: Why is WTP low in developing countries?**

- Lecture 6: Revealed preference approaches
- Lecture 7: Incentive compatible approaches

## Course Outline

**Module 5: Environmental Policy Design**

- Lecture 8: Monitoring, enforcement
- Lecture 9: Barriers to optimal design
\vfill

**Module 6: Political Economy of the Environment**

- Lecture 10: Electoral cycles, corruption
- Lecture 11: State Capacity

## Course Outline

**Module 7: Research Proposal Presentations**

- Lecture 12: TBD
- Lecture 13: Presentations
- Lecture 14: Presentations

@@

\begin{table}[]
\begin{tabular}{lllll}
In-class presentations      & 10\% &  &  &  \\
Replication Assignment           & 20\% &  &  &  \\
Research Proposal & 60 \% &  &  &  \\
Participation                     & 10\%     &  &  & 
\end{tabular}
\end{table}

@@

- I want you to become expert conference presenters after taking this class

- At start of **each** class, you'll give a 10 min paper presentation
  - The paper for presentation is on the syllabus

- Each student submits **10** summary slides (5% of grade)
  - motivation, research question, methods, results
  - 10 mins presentation + 5 mins Q&A (5% of grade)

@@

- You will replicate an environment/development paper
  - You will also **extend** the results
  - Many papers on the syllabus have replication files

- You will submit a write-up explaining what you did

@@

\begin{table}[]
\begin{tabular}{lllll}
Written Proposal      & 30\% & Oct. 31  &  &  \\
First Draft      & pass/fail \% & Oct. 3rd  &  &  \\
Peer Review           & 20\% & Oct. 10th  &  &  \\
Proposal Presentation & 10\% & Oct 15/17  &  &  \\
                      &      &  &  & 
\end{tabular}
\end{table}

- You will develop a research proposal for an original idea
  - You are NOT expected to actually do the analysis  
  - I will provide small deadlines (first draft, etc.) along the way

- Come to office hours to pitch your idea

@@

## Next Class

@@

- Impact of economic development on the environment (income effects)

@@

\huge
Aside: Best Practice for Short Presentations

## Best Practices: Structure

- You cannot present a paper in 10 minutes
  - Do not give detailed lit review or go into extreme detail

\vfill
- Instead, you are giving a trailer for the movie
  - Convince the audience that they should read the paper

\vfill
- Your goal is only to state why the paper is important, and what you did
\vfill

- Convey paper's importance in first slide

## Best Practices: Slides

- Motivation + broad research question (1 slide)

- Full paper overview (1 slide)

- Lit review (optional, 1 slide)

- Background (1 slide)

- Data (2 slides)

- Empirical Strategy (2 slides)

- Results (2 slides)

- Summary (1 slide)

## Example of Preview Slide

- **Question:** How do firms react to tribal forest policy?

- **Idea:** Model aggregate economic response and changes in firm composition

- **Setting:** India Forest Rights Act (2008)
  - Imposes transaction cost on firms

- **Data:** Manufacturing census (2001-2015); Deforestation permits (2001-2021)

- **Empirical Strategy:** Diff-in-diff using policy shift in tribal and non-tribal district

### Results Preview
1) decline in firm activity, 2) less forest encroachment by industry\
3) larger, but less productive firms survive

## Best Practices: Slide Format

- 1 minute per slide

- Avoid chart junk

- One line bullets
  - No need for full sentences

- Vary text slides and text + image slides

- Don't put too many equations
  - Save details for speaking, or talk ``about'' equation

- Summarize findings again at the end

## Best Practices: Presentation

- Speak clearly and loudly

- Speak slowly

- Look at audience; Do not show your back

- Do not stand in front of slides

- Avoid jargon

- Avoid pacing around room

- Stick to your time limit

### 2024-08-30 lectures/module2/module2.Rmd
update first week

---
title: "| Environmental and Development Economics\n| Week 2 - How does development affect the environment?
  \n"
author: |
  | Raahil Madhok
  | UMN Apllied Economics
date: "`r Sys.Date()`"
output:
  beamer_presentation:
    theme: Montpellier
  slidy_presentation: default
  ioslides_presentation: default
classoption: aspectratio=169
header-includes:
- "\\AtBeginDocument{\\title[Week 2 - development and environment]{Environmental and Development
  Economics \\\\ Week 2 - How does development affect the environment?}}"
- \input{header.tex}
---

>>>>

---
title: "| Environmental and Development Economics\n| Week 2 - How does development affect the environment?
  \n"
author: |
  | Raahil Madhok
  | UMN Apllied Economics
date: "`r Sys.Date()`"
output:
  beamer_presentation:
    theme: Montpellier
  slidy_presentation: default
  ioslides_presentation: default
classoption: aspectratio=169
header-includes:
- "\\AtBeginDocument{\\title[Module 2 - development and environment]{Environmental and Development
  Economics \\\\ Module 2 - How does development affect the environment?}}"
- \input{header.tex}
---
